---
title: "Homework_MD_06"
author: "Jessie Goldblatt"
embed-resources: true
format: html
editor: visual
---

## Setting Up

I am going to begin by loading the tidyverse package and setting my ggplot theme.

```{r}
#| message: false
library(tidyverse)
# Set ggplot theme from the outset
theme_set(theme_light())
```

I am now going to read in and look at data on the employment conditions of women across different occupations.

```{r}
#| message: false
# Read in the data in 
gender_employment <- read_csv("gender_employment.csv")

# Glimpse at the data 
glimpse(gender_employment)
```

Next, I am going to take a bird's eye view of women's earnings with respect to men's over time, using the code given in the instructions.

```{r}
gender_employment%>% 
  ggplot(aes(x = year, y = wage_percent_of_male)) +
  geom_jitter(alpha = 0.1) + 
  geom_smooth(method = "lm") + 
  labs(title = "Women's earnings with respect to men's", 
       y = "% of Men's Income", 
       x = "Year")
```

There is not a lot of fluctuation over time; the trend is mostly flat with a slight positive slope. In the following sections, I am going to explore how this relationship varies *across* *occupations*.
